The Galactic Delay-Time Distribution of Type Ia Supernovae:\\ A Chemical Evolution Perspective

### 2023-06-30 src/tex/ms.tex
[showyourwork] overleaf sync

We use APOGEE DR17 \citep{Abdurro'uf2022-SDSSIV-DR17}. Our selection criteria are presented in Table \ref{tab:sample-selection}, producing a final sample of 192,926 stars.

@@

\citet{Greggio2005-AnalyticalRates} derives analytical DTDs for SD and DD progenitor systems from assumptions about binary stellar evolution and outcomes of mass exchange. They find that the parameters which have a large effect on the shape of the DTD are the distribution and range of stellar masses in progenitor systems; the efficiency of accretion in the SD scenario; and the distribution of separations at birth in the DD scenario. Figure \ref{fig:greggio05-dtd} shows the analytical DTDs for SD progenitors and two different prescriptions for DD progenitors (``WIDE'' and ``CLOSE''). In the ``WIDE'' scheme, it is assumed that there is a wide distribution of ratios $A/A_0$ of the separation of the DD system to the initial separation of the binary, and that the distributions of $A$ and total mass of the system $m_{\rm{DD}}$ are independent, so one cannot necessarily predict the total merge time of a system based on its initial parameters. In the ``CLOSE'' scheme, there is assumed to be a narrow distribution of $A/A_0$ and a correlation between $A$ and $m_{\rm{DD}}$, so the most massive binaries tend to merge quickly and the least massive merge last.

@@

\noindent We use a non-linear least squares method to fit Equation \ref{eq:greggio05-approx} to the analytical DTDs for DD progenitors. We obtain best-fit parameters $\alpha_1=-0.19$, $A_{\rm{rise}}=1.7$, $\tau_{\rm{rise}}=0.084$ Gyr, $\alpha_2=-0.91$, $A_{\rm{tail}}=0.85$, and $\tau_{\rm{tail}}=0.088$ Gyr for the ``WIDE'' scheme, and $\alpha_1=-0.53$, $A_{\rm{rise}}=3.0$, $\tau_{\rm{rise}}=0.042$ Gyr, $\alpha_2=-1.1$, $A_{\rm{tail}}=0.62$, and $\tau_{\rm{tail}}=0.11$ Gyr for the ``CLOSE'' scheme.\footnote{
Our implementation of these best-fit approximations, as well as the original SD and DD formulations, is available in the GitHub repository for this paper.
} These best-fit approximations are represented by the dashed curves in Figure \ref{fig:greggio05-dtd} \citep[see also Figure 8 in][]{Greggio2005-AnalyticalRates}. The difference between the true analytical DTDs and our best-fit approximations is small enough as to be irrelevant for our purposes, so for the rest of this section we rely on the best-fit approximations for faster simulations.

@@

\begin{deluxetable*}{ll|cccc}
\label{tab:results}
\tablehead{
\colhead{DTD} & \colhead{SFH} & \colhead{MDF} & \colhead{ODF} & \colhead{[Fe/H]--[O/Fe]} & \colhead{Age--[O/Fe]}
}
\startdata
Power law               & Inside-out    & \meh  & \no   & \meh  & \no     \\
($\alpha=-1.1$)         & Late burst    & \meh  & \no   & \meh  & \no     \\
                        & Early burst   & \no   & \yes  & \meh  & \meh    \\
                        & Two-infall    & \no   & \no   & \meh  & \no     \\
\hline
Power law               & Inside-out    & \meh  & \no   & \no   & \no     \\
($\alpha=-1.4$)         & Late burst    & \meh  & \no   & \no   & \no     \\
                        & Early burst   & \no   & \yes  & \no   & \meh    \\
                        & Two-infall    & \no   & \no   & \no   & \no     \\
\hline
Exponential             & Inside-out    & \yes  & \no   & \yes  & \no     \\
($\tau=1.5$ Gyr)        & Late burst    & \yes  & \no   & \yes  & \no     \\
                        & Early burst   & \no   & \yes  & \meh  & \yes    \\
                        & Two-infall    & \yes  & \meh  & \meh  & \meh    \\
\hline
Exponential             & Inside-out    & \yes  & \meh  & \yes  & \meh    \\
($\tau=3$ Gyr)          & Late burst    & \yes  & \meh  & \yes  & \meh    \\
                        & Early burst   & \no   & \yes  & \no   & \yes    \\
                        & Two-infall    & \meh  & \meh  & \meh  & \meh    \\
\hline
Plateau                 & Inside-out    & \yes  & \meh  & \yes  & \no     \\
($W=0.3$ Gyr)           & Late burst    & \yes  & \no   & \yes  & \no     \\
                        & Early burst   & \no   & \yes  & \meh  & \yes    \\
                        & Two-infall    & \yes  & \meh  & \yes  & \meh    \\
\hline
Plateau                 & Inside-out    & \meh  & \meh  & \yes  & \meh    \\
($W=1.0$ Gyr)           & Late burst    & \meh  & \meh  & \yes  & \meh    \\
                        & Early burst   & \no   & \yes  & \no   & \yes    \\
                        & Two-infall    & \meh  & \meh  & \meh  & \meh    \\
\hline
Prompt                  & Inside-out    & \meh  & \no   & \no   & \no     \\
($t_{\rm max}=0.05$ Gyr)& Late burst    & \meh  & \no   & \no   & \no     \\
                        & Early burst   & \no   & \yes  & \no   & \yes    \\
                        & Two-infall    & \yes  & \no   & \no   & \no     \\
\hline
Triple system           & Inside-out    & \yes  & \meh  & \meh  & \yes    \\
($t_{\rm{max}}=1$ Gyr)  & Late burst    & \yes  & \yes  & \meh  & \meh    \\
                        & Early burst   & \no   & \meh  & \no   & \yes    \\
                        & Two-infall    & \meh  & \meh  & \meh  & \meh    \\
\enddata
\end{deluxetable*}

@@

Blah

\begin{acknowledgments}
Funding for the Sloan Digital Sky 
Survey IV has been provided by the 
Alfred P. Sloan Foundation, the U.S. 
Department of Energy Office of 
Science, and the Participating 
Institutions.

@@

SDSS-IV is managed by the 
Astrophysical Research Consortium 
for the Participating Institutions 
of the SDSS Collaboration including 
the Brazilian Participation Group, 
the Carnegie Institution for Science, 
Carnegie Mellon University, Center for 
Astrophysics | Harvard \& 
Smithsonian, the Chilean Participation 
Group, the French Participation Group, 
Instituto de Astrof\'isica de 
Canarias, The Johns Hopkins 
University, Kavli Institute for the 
Physics and Mathematics of the 
Universe (IPMU) / University of 
Tokyo, the Korean Participation Group, 
Lawrence Berkeley National Laboratory, 
Leibniz Institut f\"ur Astrophysik 
Potsdam (AIP),  Max-Planck-Institut 
f\"ur Astronomie (MPIA Heidelberg), 
Max-Planck-Institut f\"ur 
Astrophysik (MPA Garching), 
Max-Planck-Institut f\"ur 
Extraterrestrische Physik (MPE), 
National Astronomical Observatories of 
China, New Mexico State University, 
New York University, University of 
Notre Dame, Observat\'ario 
Nacional / MCTI, The Ohio State 
University, Pennsylvania State 
University, Shanghai 
Astronomical Observatory, United 
Kingdom Participation Group, 
Universidad Nacional Aut\'onoma 
de M\'exico, University of Arizona, 
University of Colorado Boulder, 
University of Oxford, University of 
Portsmouth, University of Utah, 
University of Virginia, University 
of Washington, University of 
Wisconsin, Vanderbilt University, 
and Yale University.
\end{acknowledgments}

@@

>>>>

We use APOGEE DR17 \citep{Abdurro'uf2022-SDSSIV-DR17}. Our selection criteria are presented in Table \ref{tab:sample-selection}, producing a final sample of \variable{output/sample_size.txt} stars. We limit to stars with a reported measurement of both [Fe/H] and [O/Fe]. APOGEE survey \citep{Majewski2017-APOGEE}, spectrographs \citep{Wilson2019-APOGEE-Spectrographs}, ASPCAP \citep{Holtzmann2015-ASPCAP,GarciaPerez2016-ASPCAP}, 2.5-meter Sloan Foundation Telescope \citep{Gunn2006-SloanTelescope} at Apache Point Observatory, New Mexico. Ir{\'e}n{\'e}e DuPont Telescope \citep{BowenVaughan1973-DuPontTelescope} at Las Campanas Observatory, Chile for Southern Hemisphere observations. Data reduction pipeline \citep{Nidever2015-APOGEE-DataReduction}, model grids and interpolation method \citep{Jonsson2020-APOGEE-DR16}.

@@

\citet{Greggio2005-AnalyticalRates} derives analytical DTDs for SD and DD progenitor systems from assumptions about binary stellar evolution and outcomes of mass exchange. They find that the parameters which have a large effect on the shape of the DTD are the distribution and range of stellar masses in progenitor systems; the efficiency of accretion in the SD scenario; and the distribution of separations at birth in the DD scenario. Figure \ref{fig:dtd-greggio05} shows the analytical DTDs for SD progenitors and two different prescriptions for DD progenitors (``WIDE'' and ``CLOSE''). In the ``WIDE'' scheme, it is assumed that there is a wide distribution of ratios $A/A_0$ of the separation of the DD system to the initial separation of the binary, and that the distributions of $A$ and total mass of the system $m_{\rm{DD}}$ are independent, so one cannot necessarily predict the total merge time of a system based on its initial parameters. In the ``CLOSE'' scheme, there is assumed to be a narrow distribution of $A/A_0$ and a correlation between $A$ and $m_{\rm{DD}}$, so the most massive binaries tend to merge quickly and the least massive merge last.

@@

\noindent We use a non-linear least squares method to fit Equation \ref{eq:greggio05-approx} to the analytical DTDs for DD progenitors. We obtain best-fit parameters $\alpha_1=-0.19$, $A_{\rm{rise}}=1.7$, $\tau_{\rm{rise}}=0.084$ Gyr, $\alpha_2=-0.91$, $A_{\rm{tail}}=0.85$, and $\tau_{\rm{tail}}=0.088$ Gyr for the ``WIDE'' scheme, and $\alpha_1=-0.53$, $A_{\rm{rise}}=3.0$, $\tau_{\rm{rise}}=0.042$ Gyr, $\alpha_2=-1.1$, $A_{\rm{tail}}=0.62$, and $\tau_{\rm{tail}}=0.11$ Gyr for the ``CLOSE'' scheme.\footnote{
Our implementation of these best-fit approximations, as well as the original SD and DD formulations, is available in the GitHub repository for this paper: \GitHubURL
} These best-fit approximations are represented by the dashed curves in Figure \ref{fig:dtd-greggio05} \citep[see also Figure 8 in][]{Greggio2005-AnalyticalRates}. The difference between the true analytical DTDs and our best-fit approximations is small enough as to be irrelevant for our purposes, so for the rest of this section we rely on the best-fit approximations for faster simulations.

@@

\begin{deluxetable*}{ll|ccccc}
\label{tab:results}
\tablehead{
    \colhead{DTD} & \colhead{SFH} & \colhead{[Fe/H] DF} & \colhead{[O/Fe] DF} & 
    \colhead{[Fe/H]--[O/Fe]} & \colhead{Age--[O/Fe]} & \colhead{Bimodality}
}
\startdata
Power law & Inside-out & \no & \meh & \meh & \no & \no \\
($\alpha=-1.1$) & Late-burst & \no & \no & \meh & \no & \no \\
 & Early-burst & \yes & \no & \no & \meh & \yes \\
 & Two-infall & \yes & \meh & \meh & \no & \no \\ 
\hline
Power law & Inside-out & \no & \no & \no & \no & \no \\
($\alpha=-1.4$) & Late-burst & \no & \no & \no & \no & \no \\
 & Early-burst & \meh & \no & \no & \no & \yes \\
 & Two-infall & \no & \no & \no & \no & \no \\ 
\hline
Exponential & Inside-out & \meh & \meh & \meh & \no & \no \\
($\tau=1.5$ Gyr) & Late-burst & \meh & \yes & \meh & \meh & \yes \\
 & Early-burst & \no & \no & \no & \meh & \yes \\
 & Two-infall & \yes & \meh & \meh & \meh & \yes \\ 
\hline
Exponential & Inside-out & \yes & \yes & \yes & \meh & \no \\
($\tau=3.0$ Gyr) & Late-burst & \meh & \yes & \yes & \yes & \yes \\
 & Early-burst & \meh & \no & \meh & \yes & \yes \\
 & Two-infall & \yes & \meh & \meh & \yes & \yes \\ 
\hline
Plateau & Inside-out & \meh & \yes & \yes & \meh & \no \\
($W=0.3$ Gyr) & Late-burst & \no & \yes & \yes & \meh & \yes \\
 & Early-burst & \meh & \meh & \no & \yes & \yes \\
 & Two-infall & \yes & \yes & \yes & \meh & \yes \\ 
\hline
Plateau & Inside-out & \yes & \yes & \yes & \meh & \no \\
($W=1.0$ Gyr) & Late-burst & \no & \yes & \yes & \yes & \yes \\
 & Early-burst & \meh & \meh & \meh & \yes & \yes \\
 & Two-infall & \yes & \yes & \yes & \yes & \yes \\ 
\hline
Prompt & Inside-out & \no & \no & \no & \no & \no \\
($t_{\rm max}=0.05$ Gyr) & Late-burst & \no & \no & \no & \no & \yes \\
 & Early-burst & \yes & \no & \no & \meh & \yes \\
 & Two-infall & \meh & \meh & \no & \no & \no \\ 
\hline
Triple system & Inside-out & \yes & \yes & \yes & \yes & \no \\
($t_{\rm max}=1.0$ Gyr) & Late-burst & \no & \yes & \yes & \yes & \yes \\
 & Early-burst & \meh & \meh & \meh & \yes & \yes \\
 & Two-infall & \yes & \meh & \yes & \yes & \yes \\
\enddata
\end{deluxetable*}

\subsection{Radial Migration \& Bimodality}

Something about \citet{Johnson2021-Migration} and \citet{Schonrich2009-RadialMixing}.

@@

\begin{itemize}

    \item Our diagnostics consistently favor a DTD with a large fraction of delayed SNe Ia, such as the 3 Gyr exponential, 1 Gyr plateau, or triple system evolution models, over a DTD with a high number of prompt SNe Ia, such as the $t^{-1.4}$ power law or prompt models. [Discuss the disconnect and/or complement with \citet{Maoz2017-CosmicDTD}, etc.; something about different timescales probed by volumetric surveys vs. us]

    \item While the choice of DTD cannot produce a bimodal distribution of [O/Fe], it can affect the shape of the distribution, the ratio of high-$\alpha$ to low-$\alpha$ stars, and the location of the high-$\alpha$ sequence if it exists. The SFH is the critical determining factor in bimodality.

    \item The distribution of [Fe/H] depends strongly on the chosen SFH, but there is only a weak effect from the DTD.

    \item The early-burst SFH model is the only one which produced a bimodal distribution of [O/Fe] across the disk resembling APOGEE, from the models we investigate. [Add: summary of why everything else fails.]

    \item [Something about how radial migration can't produce bimodality on its own]

\end{itemize}

\begin{acknowledgments}

Funding for the Sloan Digital Sky 
Survey IV has been provided by the 
Alfred P. Sloan Foundation, the U.S. 
Department of Energy Office of 
Science, and the Participating 
Institutions.

@@

SDSS-IV is managed by the 
Astrophysical Research Consortium 
for the Participating Institutions 
of the SDSS Collaboration including 
the Brazilian Participation Group, 
the Carnegie Institution for Science, 
Carnegie Mellon University, Center for 
Astrophysics | Harvard \& 
Smithsonian, the Chilean Participation 
Group, the French Participation Group, 
Instituto de Astrof\'isica de 
Canarias, The Johns Hopkins 
University, Kavli Institute for the 
Physics and Mathematics of the 
Universe (IPMU) / University of 
Tokyo, the Korean Participation Group, 
Lawrence Berkeley National Laboratory, 
Leibniz Institut f\"ur Astrophysik 
Potsdam (AIP),  Max-Planck-Institut 
f\"ur Astronomie (MPIA Heidelberg), 
Max-Planck-Institut f\"ur 
Astrophysik (MPA Garching), 
Max-Planck-Institut f\"ur 
Extraterrestrische Physik (MPE), 
National Astronomical Observatories of 
China, New Mexico State University, 
New York University, University of 
Notre Dame, Observat\'ario 
Nacional / MCTI, The Ohio State 
University, Pennsylvania State 
University, Shanghai 
Astronomical Observatory, United 
Kingdom Participation Group, 
Universidad Nacional Aut\'onoma 
de M\'exico, University of Arizona, 
University of Colorado Boulder, 
University of Oxford, University of 
Portsmouth, University of Utah, 
University of Virginia, University 
of Washington, University of 
Wisconsin, Vanderbilt University, 
and Yale University.

\end{acknowledgments}

@@

\section{Reproducibility}
\label{app:reproducibility}

This study was carried out using the reproducibility software
\href{https://github.com/showyourwork/showyourwork}{\showyourwork}
\citep{Luger2021-showyourwork}, which leverages continuous integration to
programmatically download the data from
\href{https://zenodo.org/}{zenodo.org}, create the figures, and
compile the manuscript. Each figure caption contains two links: one
to the dataset stored on zenodo used in the corresponding figure,
and the other to the script used to make the figure (at the commit
corresponding to the current build of the manuscript). The git
repository associated to this study is publicly available at
\GitHubURL, and the release v.X.X allows anyone to re-build the entire 
manuscript. The datasets are stored at [URL].
